---
title: "HW6"
output: github_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

```{r import-lib, message=F}
library(tidyverse)
library(modelr)
library(latex2exp)
```

## Problem 1

### Data filtering
```{r problem1-data-import}
omit_location = c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")
omit_race = c("White", "Black")

homicide = read_csv("./data/homicide-data.csv") 
homicide = homicide |>
  mutate(city_state = str_c(city, state, sep = ", "),
         resolved = as.numeric(disposition == "Closed by arrest"),
         victim_age = as.numeric(victim_age)) |>
  filter(!city_state %in% omit_location & 
         victim_race %in% omit_race &
         victim_sex != "Unknown") |>
  drop_na(victim_age) |>
  select(resolved, victim_age, victim_race, victim_sex, city_state)
```


### Model fitting
```{r}
logit_model = homicide |> 
  filter(city_state == "Baltimore, MD") |>
  glm(resolved ~ victim_age + victim_sex + victim_race, data = _, family = "binomial")
broom::tidy(logit_model, conf.int = T) |>
  mutate(OR = exp(estimate),
         OR.conf.low = exp(conf.low),
         OR.conf.high = exp(conf.high)) |>
  filter(str_detect(term, "sex")) |>
  select(term, OR, OR.conf.low, OR.conf.high) |>
  knitr::kable(digits = 3)
```


```{r}
tidy_stats = function(city, df){
  df |> 
    filter(city_state == city) |>
    glm(resolved ~ victim_age + victim_sex + victim_race, data = _, family = "binomial") |>
    broom::tidy(conf.int = T) |>
    mutate(OR = exp(estimate),
           OR.conf.low = exp(conf.low),
           OR.conf.high = exp(conf.high)) |>
    filter(str_detect(term, "sex")) |>
    select(term, OR, OR.conf.low, OR.conf.high)
}

cities = homicide |> pull(city_state) |> unique()
test_on_city = tibble(
  city = cities,
  hypo = map(cities, tidy_stats, df = homicide)
) |> unnest(hypo) |>
  select(-term)
test_on_city |>
  knitr::kable(digits = 3)
```

### Plot

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.
```{r}
test_on_city |>
  mutate(city = fct_reorder(city, OR)) |>
  ggplot(aes(x = city, y = OR, color = city)) +
  geom_point() +
  geom_errorbar(aes(ymin = OR.conf.low, ymax = OR.conf.high)) +
  theme_bw() +
  labs(x = "city", y = "OR", title = "OR of sex") +
  theme(plot.title = element_text(hjust = 0.5), axis.title.x=element_blank(),
        axis.text.x=element_blank(), axis.ticks.x=element_blank())

```



## Problem 2

### Data import
```{r problem2-data-load}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

### Bootstrapping

```{r}
mlr = function(resample){
  df = resample |> as_tibble()
  model = lm(tmax ~ tmin + prcp, data = df)
  return(
    tibble(
      residual_sq = broom::glance(model) |> pull(r.squared),
      log_beta1_beta2 = broom::tidy(model) |> 
        filter(term != "(Intercept)") |>
        pull(estimate) |> prod() |> log()
  ))
  
}

boot_straps = modelr::bootstrap(weather_df, n = 5000) |> 
  mutate(result = map(strap, mlr)) |>
  unnest(result)
```

### Plot
```{r}
boot_straps |> ggplot(aes(x = residual_sq)) +
  geom_density(fill = "steelblue", alpha = .7, color = "steelblue") +
  theme_bw() +
  labs(x = TeX("$r^2$"), y = "density", title = TeX("Distribution of $r^2$")) +
  theme(plot.title = element_text(hjust = 0.5))
```

Description: The distribution of $r^2$ is approximately symmetric. The mean is around 0.92.

```{r}
boot_straps |> ggplot(aes(x = log_beta1_beta2)) +
  geom_density(fill = "tomato", alpha = .7, color = "tomato") +
  theme_bw() +
  labs(x = TeX("$\\log(\\beta_1*\\beta_2)$"), y = "density", title = TeX("Distribution of $\\log(\\beta_1*\\beta_2)$")) +
  theme(plot.title = element_text(hjust = 0.5))

```

Description: All values are less than 0. The distribution is left-skewed, which may be caused by the `NA` generated during the logarithm calculation.
The `NA` ratio is `r boot_straps |> pull(log_beta1_beta2) |> is.na() |> sum() / dim(boot_straps)[1]`.

### Quantile for CI
```{r}
ci = function(data){
  quantile(data, na.rm = T, probs = c(0.025, 0.975)) |> round(digits = 3)
}
```

The 95% confidence interval for $r^2$ is (`r ci(boot_straps$residual_sq)[1]`, `r ci(boot_straps$residual_sq)[2]`).

The 95% confidence interval for $\log(\beta_1*\beta_2)$ is (`r ci(boot_straps$log_beta1_beta2)[1]`, `r ci(boot_straps$log_beta1_beta2)[2]`).

## Problem 3

#### Preprocessing
```{r}
birth_weight = read.csv("./data/birthweight.csv") |> tibble()
birth_weight = birth_weight |>
  mutate(babysex = factor(recode(babysex, `1` = "male", `2` = "female")),
         frace = factor(recode(frace, `1` = "white",  `2` = "black", `3` = "asian", `4` = "puerto rican",`8` = "other", `9` = "unknown")),
         mrace = factor(recode(mrace, `1` = "white", `2` = "black", `3` = "asian", `4` = "puerto rican", `8` = "other")),
         malform = factor(recode(malform, `0` = "absent", `1` = "present"))) |>
  select(-pnumlbw, -pnumsga)

```

Data cleaning:
- Mutate `babysex`, `frace`, `mrace` and `malform` as factor because they are categorical variables.
- Remove `pnumlbw` and `pnumsga` because they both only 0 in all values, indicating they are useless variables.


#### Model description
```{r}
# model with full variables
mlr = lm(bwt ~ ., data = birth_weight)
mlr |> summary()
```
By using all variables for linear regression, I found that `babysex`, `bhead`, `blength`, `delwt`, `gaweeks`,  `parity`, `smoken` are significant variables and coefficient of `wtgain` variable is fitted as `NA`(maybe caused by collinearity), while other variables are not significant.


#### Fitting MLR & Plot
```{r}
my_mlr = lm(bwt ~ ., data = birth_weight)
broom::tidy(my_mlr) |>
  knitr::kable(digits = 3)
```

```{r}
birth_weight |> 
  modelr::add_predictions(my_mlr) |>
  modelr::add_residuals(my_mlr) |>
  select(pred, resid) |>
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .5) + 
  theme_bw() +
  labs(x = "Fitted value", y = "Residual", title = "Residual-Fitted value") +
  theme(plot.title = element_text(hjust = 0.5))
```


#### Model comparison with cross validation

Compare model to two others:
One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

```{r}
model_eval = function(train_df, test_df){
  my_mlr = lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + parity + smoken, data = train_df)
  mlr1 = lm(bwt ~ blength + gaweeks, data = train_df)
  mlr2 = lm(bwt ~ bhead * blength * babysex, data = train_df)
  return(
    tibble(
      rmse0 = rmse(my_mlr, test_df), 
      rmse1 = rmse(mlr1, test_df), 
      rmse2 = rmse(mlr2, test_df)
    )
  )
}
```

```{r}
times = 100
rmse_df = birth_weight |>
  modelr::crossv_mc(n = times) |>
  mutate(train = map(train, as.tibble), 
         test = map(test, as.tibble), 
         eval_mse = map2(train, test, model_eval)) |>
  unnest(eval_mse) |>
  select(starts_with("rmse"))
```

```{r}
rmse_df |>
  pivot_longer(cols = starts_with("rmse"), names_to = "model_slc", values_to = "rmse",  names_prefix = "rmse") |>
  mutate(model = recode(model_slc, 
                        "0" = "my model", 
                        "1" = "model 1", 
                        "2" = "model 2")) |>
  ggplot(aes(x = rmse, fill = model)) +
  geom_density(alpha = .4, adjust = .8, color = "steelblue") +
  labs(
    title = "Density of RMSE in different model"
  ) + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5))
```



